---
title: "10 - Running through H2O Documentation"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(h2o)
h2o.init()
```

# Getting Data into your H2O Cluster
Here is the process for [uploading a file](http://h2o-release.s3.amazonaws.com/h2o/rel-yau/5/docs-website/h2o-docs/data-munging/uploading-data.html) and [importing a file](http://h2o-release.s3.amazonaws.com/h2o/rel-yau/5/docs-website/h2o-docs/data-munging/importing-data.html).

I am starting with a .csv created from the [Clean-LAPOP repository]() that has years 2008-2014 of factored data for Argentina.
```{r}
#Selects file
argentina_path <- "argentina copy.csv"

#Uploads to H2O
argentina_factored <- h2o.uploadFile(path = argentina_path, destination_frame = "argentina_factored")
```

# Combining Columns from Two Datasets
Using the example from the documentation to create simple data frames.

```{r}
left <- data.frame(fruit = c("apple", "orange", "banana"), color = c("red", "orange", "yellow"))

right <- data.frame(fruit = c("apple", "orange", "banana"), citrus = c(FALSE, TRUE, FALSE))

#Create the H2O data frames from the inputted data
l.hex <- as.h2o(left)
l.hex

r.hex <- as.h2o(right)
r.hex

#Combine the two datasets into single dataset
columns.hex <- h2o.cbind(l.hex, r.hex)
columns.hex
```

From this I learned that H2O uses "0" to designate the overlapping column. I wonder if you are able to use Tidyverse with H2O.

# Combining Rows from Two Datasets
Will combine rows using example data from website.

```{r}
#Creates path to file
ecg1Path <- "http://h2o-public-test-data.s3.amazonaws.com/smalldata/anomaly/ecg_discord_train.csv"

#Makes it an object
ecg1.hex <- h2o.importFile(path=ecg1Path, destination_frame="ecg1.hex")

#Prints dimensions
print(dim(ecg1.hex))

#Same as above
ecg2Path <- "http://h2o-public-test-data.s3.amazonaws.com/smalldata/anomaly/ecg_discord_test.csv"

ecg2.hex <- h2o.importFile(path=ecg2Path, destination_frame="ecg2.hex")

print(dim(ecg2.hex))

#Combine into single dataset
ecgCombine.hex <- h2o.rbind(ecg1.hex, ecg2.hex)
print(dim(ecgCombine.hex))
```

This seems to be pretty similar to simple combinations of datasets in base R.

# Fill NAs
This function fills in NA values in a sequential manner. Need to specify if it goes forward or backward, whether it should go by rows or columns, and max number of consecutive NAs to fill.

Using example from documentation.
```{r}
#Creates random data frame where no more than 70% of values are NAs.
fr_with_nas <- h2o.createFrame(categorical_fraction = 0.0,
                               missing_fraction = 0.7,
                               rows = 6,
                               cols = 2,
                               seed = 123)
fr_with_nas
```

I am going to pause here with notes on Data Manipulation section, will refer to the documentation if needed. Particularly because I feel that I may be able to use Tidyverse in conjunction with this package.