---
title: "11 - Reading MDSR Chapter 7"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Chapter 7 - Statistical Foundations

```{r, message=FALSE}
library(mdsr)
library(tidyverse)
library(nycflights13)
```

# 7.1
## Example: Sampling from the population
A traveler has a meeting in San Francisco (SFO) at time **t**. Try to find how much earlier than **t** an acceptable flight shoudl arrive in order to avoid being late to the meeting due to a flight delay.

```{r}
View(flights)
str(flights)
```


Creating the population for this example - all flights that arrived in SFO in 2013.
```{r}
SF <- flights %>% 
  filter(dest == "SFO", !is.na(arr_delay))

View(SF)
```

Setting sample size to n = 25 cases.
```{r}
set.seed(101)
Sample25 <- SF %>% 
  sample_n(size = 25)
```

Looking at results from the sample
```{r}
favstats(data = Sample25, ~ arr_delay)
```

Comparing to population
```{r}
favstats(data = SF, ~ arr_delay)
```

98th percentile of the arrival delays in our data sample:
```{r}
qdata(~ arr_delay, p = 0.98, data = Sample25)
```

Testing to see if this policy would have worked by using the population data
```{r}
tally(~ arr_delay < 90, data = SF, format = "proportion")
```
This shows that our policy misses the mark 5% of the time - which is worse than intended. We need to increase it from 90 minutes but to what??

Since we have the population, you can calculate the 98th percentile of arrival delays
```{r}
qdata(~ arr_delay, p = 0.98, data = SF)
```
Now we know that it should have been about 150 minutes.

In the future when we don't have access to population, how can we find the 98th percentile with *only* the sample?


# 7.2 Sample Statistics
A more reliable sample statistic would be using the median. We ultimately want to know how well the sample statistic reflects the population.

## The sampling distribution
IF we draw many different samples from the population, each of size n, and calculated the sample statistic on each of those samples, how similar would the sample statistic be across all the samples?

```{r}
n <- 25
mean(~ arr_delay, data = sample_n(SF, size = n, replace = FALSE))
```

Now we need to repeat it multiple times.
```{r}
trials <- do(500) *
  mean(~ arr_delay, data = sample_n(SF, size = n, replace = FALSE))

head(trials)
```

Now that we have the average for 500 trails, let's examine how spread out the results are.
```{r}
favstats(~ mean, data = trials)
```
Here is the same thing with large sample size - n = 100.
```{r}
trials_100 <- do(500) *
  mean(~ arr_delay, data = SF %>% 
         sample_n(size = 100, replace = FALSE))
```

Histograms of both sample sizes
```{r}
rbind(trials %>% 
        mutate(n = 25),
      trials_100 %>% 
        mutate(n = 100)) %>% 
  ggplot(aes(x = mean)) +
  geom_histogram(bins = 30) +
  facet_grid(~ n) +
  xlab("Sample mean")
```

* Larger sample sizes produce a standard error that is smaller, and it is therefore more reliable.
* Larger sample sizes tend to result in a bell-shaped distribution

# 7.3 The bootstrap
*The bootstrap* is a statistical method that allows us to approximate the sampling distribution even without access to the population.

With bootstrap, we think of the sample as the population. Through *resampling* we will draw a new sample from an existing sample but **with replacement**.

```{r}
small <- sample_n(SF, size = 3, replace = TRUE)
```

Let's use bootstrapping to find the reliability of the mean arrival time calculated on a sample size of 200. 
```{r}
n <- 200
original_sample <- SF %>% 
  sample_n(size = n, replace = FALSE)
```

From this sample, we'll draw all of the resamples and calculate the mean arrival delay.
```{r}
bootstrap_trials <- do(500) *
  mean(~ arr_delay, data = sample_n(original_sample, size = n, replace = TRUE))
favstats(~ mean, data = bootstrap_trials)
```

We can calculate the 98th percentile from our sample of size n = 100 flights, and use bootstrapping to see how reliable that sample statistic is.

```{r}
qdata(~ arr_delay, p = 0.98, data = original_sample)
```

We can check the reliability of that estimate using bootstrapping.
```{r}
bootstrap_trials <- do(500) * 
  qdata(~ arr_delay, p = 0.98,
        data = sample_n(original_sample, size = n, replace = TRUE))

favstats(~ quantile, data = bootstrap_trials)
```

The bootstrapped standard error is about 18 minutes. This would be unlikely to hit the target, and in this case even increaing the sample to 10,000 cases.

7.4 Outliers
One place where more data is helpful is in identifying outliers. Suppose we consider any flight delayed by 420 minutes (7 hours) or more as extreme.

```{r}
SF %>% 
  filter(arr_delay >= 420) %>% 
  select(month, day, dep_delay, arr_delay, carrier)
```

Most of the very long delays were in July, and Virgin America (VX) is the most frequent offender. Immediately this suggests one possible route for impoving the outcome of the business travel policy - we could tell people to arrive extra early in July and to avoid VX.

However, these outliers can be misleading, as they only account for a small fraction of flights into SFO in 2013.

*Outliers should never be dropped unless tehre is a clear rationale.*

*If outliers are dropped, this should be clearly reported.*

Histogram without the outliers.
```{r}
SF %>% 
  mutate(long_delay = arr_delay > 60) %>% 
  tally(~ long_delay | month, data = .)
```

The large majority of flights arrive without any delay or a delay less than 60 minutes. We might be able to identify patterns for when the longer delays are likely to occur.

The 14 outliers suggested that month or carrier may be linked to long delays. Let's see how that plays out with the large majority of data.

```{r}
SF %>% 
  mutate(long_delay = arr_delay > 60) %>% 
  tally(~ long_delay | month, data = .)
```

We can see that June and July are problem months.

```{r}
SF %>% 
  filter(arr_delay < 420) %>% 
  ggplot(aes(arr_delay)) +
  geom_histogram(binwidth = 15)
```

Distribution of flight arrival delays in 2013 for flights to SF from NYC airports that were delayed less than seven hours. The distribution features a long right tail (even after getting rid of the outliers).

```{r}
SF %>% 
  mutate(long_delay = arr_delay > 60) %>%
  tally(~ long_delay | carrier, data = .)
```

Delta Airines has reasonable performance. These analyses hint that a policy might advise travelers to plan to arrive extra early in June and July and to consider Delta as an airline for travel to SFO.

7.5 Statistical models: Explaining variation


