---
title: "12 - Following Gentle Intro to Tidymodels"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Notes from Gentle Intro to Tidymodels

The article can be found [here](https://rviews.rstudio.com/2019/06/19/a-gentle-intro-to-tidymodels/).

Tidymodels is a group of packages that doesn't implement the statistical models themselves, but simplifies data pre-processing and results validation.

# Four tidymodels packages sampled by modeling step:

## Pre-Process
* `rsample` - different types of re-samples
* `recipes` - transformations for model data pre-processing

## Train
* `parnip` - a common interface for model creation

## Validate
* `yardstick` - measure model performance

# Iris Example
Loading in `tidymodels` also loads in `dplyr` and `ggplot` along with other packages from `tidyverse`, so in this case `tidymodels` is the only package we need to load.

```{r}
library(tidymodels)
```

## Pre-Process
Pre-processing is where you would clean the data to prepare it for modeling. Most transformations can be completed with `tidyverse`.

## Data Sampling
`initial_split()` creates two datasets:
1. Training set
  - 3/4 of data held here
2: Testing set
  - 1/4 of data used for testing
This then creates an `rplit` object.

```{r}
iris_split <- initial_split(iris, prop = 0.6) 
#prop argument changes the proportion of data in each set
iris_split #Results: row count for testing/training/total
```

