---
title: "12 - Following Gentle Intro to Tidymodels"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Notes from Gentle Intro to Tidymodels

These notes are for personal reference only, the content comes from Gentle Intro to Tidymodels by Edgar Ruiz. The article can be found [here](https://rviews.rstudio.com/2019/06/19/a-gentle-intro-to-tidymodels/).

Tidymodels is a group of packages that doesn't implement the statistical models themselves, but simplifies data pre-processing and results validation.

# Four tidymodels packages sampled by modeling step:

## Pre-Process
* `rsample` - different types of re-samples
* `recipes` - transformations for model data pre-processing

## Train
* `parnip` - a common interface for model creation

## Validate
* `yardstick` - measure model performance

# Iris Example
Loading in `tidymodels` also loads in `dplyr` and `ggplot` along with other packages from `tidyverse`, so in this case `tidymodels` is the only package we need to load.

```{r}
library(tidymodels)
```

## Pre-Process
Pre-processing is where you would clean the data to prepare it for modeling. Most transformations can be completed with `tidyverse`.

## Data Sampling
`initial_split()` creates two datasets:
1. Training set
  - 3/4 of data held here
2: Testing set
  - 1/4 of data used for testing
This then creates an `rplit` object.

```{r}
iris_split <- initial_split(iris, prop = 0.6) 
#prop argument changes the proportion of data in each set
iris_split #Results: row count for testing/training/total
```

`training()` and `testing()` access observations from their respective datasets.

```{r}
#Viewing data from training set
iris_split %>% 
  training() %>% 
  glimpse()
```

## Pre-process interface
`recipes` package provides a data pre-processing interface.

* `recipe()` - starts a new set of transformations to be applied. Main argument is the model's formula. (Similar to `ggplot()`)
* `prep()` - executes the transformations on top of the data that is supplied (the training data, most typically)

*Each data transformation is a step* and functions correspond to specific steps.

* `step_corr()` - removes variables that have large absolute correlations with other variables
* `step_center()` - normalizes numeric data to have a mean of zero
* `step_scale()` - normalizes numeric data to have an SD of 1

These can be used in conjunction with `all_outcomes()` and `all_predictors()` to group variables.

```{r}
iris_recipe <- training(iris_split) %>% 
  recipe(Species ~.) %>% 
  step_corr(all_predictors()) %>% 
  step_center(all_predictors(), -all_outcomes()) %>% 
  step_scale(all_predictors(), -all_outcomes()) %>% 
  prep()

#Calling object prints out details about the recipe
iris_recipe
```

## Execute the pre-processing
To apply the same transformations on the testing data, we use `bake()`.

```{r}
iris_testing <- iris_recipe %>% 
  #Used to extract the right dataset
  bake(testing(iris_split))

#View results
glimpse(iris_testing)
```

To load the training data into an object, use `juice()`.
```{r}
iris_training <- juice(iris_recipe)

glimpse(iris_training)
```

## Model Training
Multiple packages can fit the same type of model and each provides a unique interface. So it is not easy to switch between packages when running the same model.

*`tidymodels` replaces the interface and provides a single set of functions and arguments to define a model then fits the model against the requested modeling package*

```{r}
iris_ranger <- rand_forest(trees = 100, mode = "classification") %>% 
  #Uses `ranger` package
  set_engine("ranger") %>% 
  fit(Species ~ ., data = iris_training)

iris_ranger
```

Using `tidymodels` you can run the same model against `randomForest`

```{r}
iris_rf <- rand_forest(trees = 100, mode = "classification") %>% 
  set_engine("randomForest") %>% 
  fit(Species ~ ., data = iris_training)

iris_rf
```

