---
title: "20 - Reading in Data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Overall Goals
Due to the resignation of the Bolivian president, Evo Morales, in October, I am interested in determining the typical supporters of his party, **Movimiento al Socialismo** (MAS), and how participants' support has changed over the past 14 years. In my project, I hope to develop a predictive model that uncovers factors associated with participant’s perception of Evo Morales and satisfaction with the Bolivian government. 

I am also going to use these models to provide a risk-adjusted measure of perception to directly answer whether residents in Bolivia feel more or less satisfied with Morales' administration than expected. I will then discern other factors that might be driving these perceptions that may not be accounted for in the survey data, and hypothesize how factors might be identified if the model predictions improve with the inclusion of new data.


## Source
In this file, we read in the data from the [LAPOP Datasets](http://datasets.americasbarometer.org/database/index.php). In order to do this, you need to download them to your machine and place them in this file.

For this section, I will be referencing and modifying code I wrote as an intern at the Vanderbilt Data Science Institute with Lindsay Hardy and Jesse Spencer-Smith; the Github repository can be found [here](https://github.com/vanderbilt-data-science/lapop-dashboard).

### Factored vs. Unfactored
The data in the Stata files are factored, meaning that a yes or no question would be entered in as 0 and 1. These numbers are then assigned labels as such: 0 - "yes" and 1 - "no", and we refer to data frames where the labels are visible in the observations as *factored*. When data frames use only the numbered observations, we refer to this as *unfactored*.

We chose to use both versions of the data frames for ease of use in Tableau Public. Unfactored data frames allow us to perform calculations in Tableau Public later on, and the factored data frames provide reader-friendly text to the graphics.

## Libraries
These packages contain the functions that are necessary to run the code below. These can be installed using `install.packages()`, and then you should run the code chunk below so they are ready to be used.
```{r, message=FALSE}
library(labelled)
library(sjmisc)
library(tidyverse)
library(assertr)
library(haven)
```

## Creating factored and unfactored dataframes
The first function we wrote, `factor_country()`, reads the file in and edits the column name so that it ends in "_factor" to differentiate between unfactored and factored when we combine them later. This will only be used once, and then we will run `read_dta()` to read in the unfactored .dta file directly from Stata.

```{r}
factor_country <- function(country){
  #Reads file for the country
  country <- read_dta(country)
  #Factors all the columns
  country <- as_factor(country)
  
  #Starts list x
  x <- 1
  
  #Changes name of column to add "factor"
  while(x <= length(country)){
      colnames(country)[x] <- paste(colnames(country)[x], "factor", sep = "_")
     x <- x + 1
  }
  country
}
```

## Bolivia Merged File
I am using the **most recent** merged country file. Currently, there is no public file with 2016 and 2018 waves merged for Bolivia.

```{r}
bolivia_factor <- factor_country("data/merged_bolivia_2004_2014.zip")
bolivia_unfactor <- read_dta("data/merged_bolivia_2004_2014.zip")
```

## Reading in 2016 file
Because there is no pre-existing merge that includes 2016 data, we are reading it in separately and then joining it later in the process.

If you are using a file that has this wave included, you can skip this.

## Creating factored data frame of 2016 questionnaire using `factor16_country`
This is the 2016 version of the `factor_country()` used above. The same process applies, but we have renamed the `uniq_id` column to `person_id` for clarification.
```{r}
factor16_country <- function(country){
  #Reads file for the country
  country <- read_dta(country)
  #Factors all the columns
  country <- as_factor(country) %>% 
    rename(person_id = uniq_id)
  
  #Starts list x
  x <- 1
  
  #Changes name of column to add "factor"
  while(x <= length(country)){
      colnames(country)[x] <- paste(colnames(country)[x], "factor", sep = "_")
     x <- x + 1
  }
  country
}
```

## Bolivia 2016 File
```{r}
bolivia_16_factor  <- factor16_country("data/2016_bolivia.dta") %>% 
  assert(is_uniq, person_id_factor)

bolivia_16_unfactor  <- read_dta("data/2016_bolivia.dta") %>% 
  rename(person_id = uniq_id) %>% 
  assert(is_uniq, person_id) %>% 
  transform(person_id = as.character(person_id))
```

## Creating factored data frame of questionnarie using `factor18_country()`
This is similar to 2016 version above; the same process applies.

```{r}
factor18_country <- function(country){
  #Reads file for the country
  country <- read_dta(country)
  #Factors all the columns
  country <- as_factor(country) %>% 
    rename(person_id = uniq_id)
  
  #Starts list x
  x <- 1
  
  #Changes name of column to add "factor"
  while(x <= length(country)){
      colnames(country)[x] <- paste(colnames(country)[x], "factor", sep = "_")
     x <- x + 1
  }
  country
}
```

## Bolivia 2018 file
```{r}
bolivia_18_factor <- factor18_country("data/2018_bolivia.dta") %>% 
  assert(is_uniq, person_id_factor)

bolivia_18_unfactor <- read_dta("data/2018_bolivia.dta") %>%
  rename(person_id = uniq_id) %>% 
  assert(is_uniq, person_id) %>% 
  transform(person_id = as.character(person_id))
```

