---
title: "25 - Data Exploration"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
I am now going to explore the tidy version of the 2004-2018 Bolivia data to find factors that influence my question: who is likely to be asked by the police for a bribe in Bolivia?

## Reading through questions
In order to familiarize myself with the questions asked by LAPOP, I'll be using `tabyl()`. This will help give me insight in what qyestions and factors may be relevant.

## Making space for memory
```{r}
rm.all.but("bolivia_clean")
```

I had to omit 2 variables, but they were unrelated to the topic, so it works out. My computer otherwise would not have had sufficient memory to widen the data.

```{r}
bolivia_clean %>% 
  #Getting questions only
  tabyl(question_es)
```

```{r}
bolivia_clean %>% 
  #Filtering to see if questions about political ideology were asked
  filter(question_es == "Ideología (izquierda / derecha)")
```

## Widening factored data
I am widening the data so that it is in the proper form for analysis with the `rpart` package.
```{r}
labelled_wide <- bolivia_clean %>% 
  #Gets first 2.2 million rows to reduce memory
  head(2200000) %>% 
  #Widening data frame
  pivot_wider(id_cols = c(person_id_factor, country, wave), 
              names_from = question_es,
              values_from = answer_dimension)
```

## Cleaning column names
```{r}
labelled_wide <- labelled_wide %>% 
  #Renaming these columns due to their length
  rename(ideology = "Ideología (izquierda / derecha)") %>% 
  rename(police_bribe = "Oficial de Policía le solicitó un soborno") %>%
  rename(perceived_safety = "Percepción de inseguridad en el barrio") %>% 
  #Cleaning names for all columns
  clean_names()
```


## Changing to NA values
```{r}
#Function to create NA values in Bolivia dataset
turn_to_NA <- function(country){
  #Setting items in list to NA
  country <- set_na(country, na = c("No se le preguntó en este país o año", "Not asked in this country or year","No Aplica", "No Sabe", "No sabe", "No Responde", "No responde"), as.tag = TRUE)
}

#Changing above values to NA
labelled_wide <- turn_to_NA(labelled_wide)

#Checking to make sure the function worked
labelled_wide %>% 
  select(police_bribe) %>% 
  filter(is.na(police_bribe))
```

# !DO NOT USE REFERENCE ONLY
#Splitting up the age values for easier reading. Ages range from 18 to 85.
#```{r}
#Bucketing values into bins
#labelled_wide <- labelled_wide %>% 
#  transform(edad = as.numeric(edad))

#labelled_wide <- labelled_wide %>% 
#  mutate(age = cut(labelled_wide$edad,
                  breaks = seq(18, 87, 10),
                  include.lowest = TRUE))

#labelled_wide
#```

#Removing this variable because it is not necessary
#```{r}
#labelled_wide <-  labelled_wide %>% 
#  select(-person_id_factor)
#```


#Wide numeric version
#```{r}
#numeric_wide <- bolivia_clean %>% 
#  head(2200000) %>% # Get first 100K questions to reduce memory %>%
#  pivot_wider(id_cols = c(person_id_factor, country, wave), 
#              names_from = question_es,
#              values_from = answer_measure)
#```