---
title: "30 - MDSR Decision Trees"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
Working to make a decision tree for the Bolivian data.

```{r}
library(mdsr)
library(rpart)
library(partykit)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
```

# Separating into Testing and Training
I will have a sample of 80% of the rows as the training dataset, and 20% for the testing dataset.

```{r}
#Setting the seed
set.seed(364)

#Getting the number of rows
n <- nrow(labelled_wide)
#Getting .2 of `labelled_wide`
test <- sample.int(n, size = round(0.2 * n))
#Making training dataset
train <- labelled_wide[-test, ]
#Checking length
nrow(train)

#Making testing dataset
test <- labelled_wide[test, ]
#Checking length
nrow(test)

#Making sure the training and testing datasets add up to `labelled_wide`
nrow(labelled_wide)
nrow(train) + nrow(test)
```

```{r}
tally(~ police_bribe, data = train)
```

I am going to try to find a split for ideology based on one variable first, urban vs. rural, and then expand.

Below, I am looking to see what the likelihood of being asked for a bribe from a police officer is based on living in an urban or rural location.
```{r}
model_one <- rpart(police_bribe ~ urbano_rural, data = train,
      method = "class",
      minsplit = 1,
      minbucket = 1,
      cp = -1)

plot(as.party(model_one))
fancyRpartPlot(model_one, main = "Individuals Asked for a Bribe by the Police in Bolivia", caption = "Source: LAPOP")
```

Now I am testing to see if there is a difference based on gender.
```{r}
model_two <- rpart(police_bribe ~ urbano_rural + sexo, data = train,
      method = "class",
      minsplit = 1,
      minbucket = 1,
      cp = -0.001)

plot(as.party(model_two))
fancyRpartPlot(model_two, main = "Individuals Asked for a Bribe by the Police in Bolivia", caption = "Source: LAPOP")
```

```{r}
model_three <- rpart(police_bribe ~ urbano_rural + sexo + perceived_safety, data = train,
      method = "class",
      minsplit = 1,
      minbucket = 1,
      cp = -1)

plot(as.party(model_three))
fancyRpartPlot(model_three, main = "Individuals Asked for a Bribe by the Police in Bolivia", caption = "Source: LAPOP")
```

```{r}
test <- rpart(police_bribe ~ urbano_rural + sexo + edad, data = train,
      method = "class",
      minsplit = 1,
      minbucket = 1,
      cp = 0)

plot(as.party(test))
fancyRpartPlot(test, main = "Individuals Asked for a Bribe by the Police in Bolivia", caption = "Source: LAPOP")
```

This doesn't work - I think it is getting overwhelmed
```{r}
model_four <- rpart(police_bribe ~ ., data = train,
      method = "class",
      minsplit = 1,
      minbucket = 1,
      cp = -0.001)

plot(as.party(model_four))
```



Jesse, the figure is best viewed in its own window. I am now looking for a way to make it easier to read what ages are in each `edad` node.

#```{r}
#formula <- as.formula("police_bribe ~ .")
#mod_tree <- rpart(formula,
#                  data = train,
#                  method = "class",
#                  minsplit = 1,
#                  minbucket = 1,
#                  cp = -1)
#mod_tree
#```

