---
title: "30 - MDSR Decision Trees"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
I am going to use a decision tree to model the typical demographic of people asked for a bribe in Bolivia. I have used Chapter 8 of the textbook *Modern Data Science for R* by Benjamin S. Baumer, et al to learn how to create statistical models. 

# Separating into Testing and Training
I will have a sample of 80% of the rows as the training dataset, and 20% for the testing dataset.
```{r}
#Setting the seed
set.seed(364)

#Getting the number of rows
n <- nrow(labelled_wide)
#Getting .2 of `labelled_wide`
test <- sample.int(n, size = round(0.2 * n))
#Making training dataset
train <- labelled_wide[-test, ]
#Checking length
nrow(train)

#Making testing dataset
test <- labelled_wide[test, ]
#Checking length
nrow(test)

#Making sure the training and testing datasets add up to `labelled_wide`
nrow(labelled_wide)
nrow(train) + nrow(test)
```

```{r}
#Checking count of answers
tally(~ police_bribe, data = train)
```

I am going to try to find a split for `police_bribe` based on one variable first, urban vs. rural, and then expand.

# Model One
Below, I am looking to see what the likelihood of being asked for a bribe from a police officer is based on living in an urban or rural location.
```{r}
model_one <- rpart(police_bribe ~ urbano_rural, data = train,
      method = "class",
      minsplit = 1,
      minbucket = 1,
      cp = -1)

plot(as.party(model_one))
fancyRpartPlot(model_one, main = "Individuals Asked for a Bribe by the Police in Bolivia", caption = "Source: LAPOP")
```

# Model Two
Now I am testing to see if there is a difference based on gender.
```{r}
model_two <- rpart(police_bribe ~ urbano_rural + sexo, data = train,
      method = "class",
      minsplit = 1,
      minbucket = 1,
      cp = -0.001)

plot(as.party(model_two))
fancyRpartPlot(model_two, main = "Individuals Asked for a Bribe by the Police in Bolivia", caption = "Source: LAPOP")
```

# Model Three
```{r}
model_three <- rpart(police_bribe ~ urbano_rural + sexo + perceived_safety, data = train,
      method = "class",
      minsplit = 1,
      minbucket = 1,
      cp = -1)

plot(as.party(model_three))
fancyRpartPlot(model_three, main = "Individuals Asked for a Bribe by the Police in Bolivia", caption = "Source: LAPOP")
```


# !DO NOT USE REFERENCE ONLY
#```{r}
#test <- rpart(police_bribe ~ urbano_rural + sexo + edad, data = train,
#      method = "class",
#      minsplit = 1,
#      minbucket = 1,
#      cp = 0)
#
#plot(as.party(test))
#fancyRpartPlot(test, main = "Individuals Asked for a Bribe by the Police in Bolivia", caption = "Source: LAPOP")
#```

#```{r}
#formula <- as.formula("police_bribe ~ .")
#mod_tree <- rpart(formula,
#                  data = train,
#                  method = "class",
#                  minsplit = 1,
#                  minbucket = 1,
#                  cp = -1)
#mod_tree
#```

