---
title: "30 - Starting Tidymodel"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
This is the first go at making a decision tree.

I am referencing the article [A Gentle Introduction to tidymodels](https://rviews.rstudio.com/2019/06/19/a-gentle-intro-to-tidymodels/) to analyze the data.

# Data Sampling
```{r, messages=FALSE}
library(tidymodels)
```

```{r}
bolivia_split <- initial_split(labelled_wide, prop = 0.6)

bolivia_split
```

Viewing the observations in the training dataset
```{r}
bolivia_split %>% 
  training() %>% 
  glimpse()
```

# Pre-process interface
```{r}
bribe_recipe <- training(bolivia_split) %>% 
  #Making a recipe object
  recipe(police_bribe ~ .) 

bribe_recipe
```

Imputing K Nearest Neighbors
```{r}
imputed <- bribe_recipe %>% 
  step_knnimpute(all_predictors())
```

Creating numeric dummy variables for categorical variables
```{r}
ind_vars <- imputed %>% 
  step_dummy(all_predictors(), -all_numeric())

ind_vars
```

All of the predictors are now numeric, so I will now center and scale them.
```{r}
standardized <- ind_vars %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors())

standardized
```

Using prep function to estimate the means and standard deviations from the training set.
```{r}
trained_recipe <- prep(standardized, training = training(bolivia_split))
```


# Execute the pre-processing
```{r}
_testing <- ideology_recipe %>% 
  bake(testing(bolivia_split))
```



